---
output: pdf_document
classoption: landscape
---

```{r set up, warning=FALSE, message=FALSE, echo = FALSE}

# Purpose: this markdown complies, annotates and formats the figures from https://github.com/JGCRI/hector-SA-npar/issues/7#event-1509615854. More details can be found there.


# Load the libs
library(dplyr)
library(ggplot2)
library(gridExtra)

# Set up the dirs
BASE    <- "C:/Users/dorh012/Documents/GitHub/CMS/hector-SA-npar"
FIG_DIR <- file.path(BASE, "diag-out")

# Load the figures
fig_list <- get(load(file.path(FIG_DIR, "figs_rcp26_UIUC.rda")))

# idk why but I was getting some errors
FIGURE_THEME <- theme()

# The color pallate to use  
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")


```


### Fig 1: "all 50,00 Hector runs - mean and spread for temperature"

```{r, fig.width=8, fig.height=8, echo = FALSE, warning=FALSE}
fig_list$all_temp
```


### Fig 2: "temperature observations on top of Hector runs"
```{r, fig.width=8, fig.height=8, echo = FALSE, warning=FALSE}
fig_list$temp_obs
```


### Fig 3: "atmo observations on top of Hector runs"
```{r, fig.width=8, fig.height=8, echo = FALSE, warning=FALSE}
fig_list$growth_obs
```


### Fig 4: "land flux observations on top of Hector runs"
```{r, fig.width=8, fig.height=8, echo = FALSE, warning=FALSE}
fig_list$landFlux_obs
```


### Fig 5: "bar chart of hector runs with filter counts"
```{r, fig.width=8, fig.height=8, echo = FALSE, warning=FALSE}
fig_list$passing_hector_runs
```

\newpage

### Fig 6: "parameter vs temperature or parameter vs CO2 plots"
#### Fig 6.A: beta 
```{r, fig.width=6, fig.height=5.5, echo = FALSE, warning=FALSE}
grid.arrange(fig_list$parameter_2100value$beta_Tgav, fig_list$parameter_2100value$beta_Ca)
```

\newpage

#### Fig 6.B: diff 
```{r, fig.width=6, fig.height=6, echo = FALSE, warning=FALSE}
grid.arrange(fig_list$parameter_2100value$diff_Tgav, fig_list$parameter_2100value$diff_Ca)
```

\newpage

#### Fig 6.C: q10
```{r, fig.width=6, fig.height=6, echo = FALSE, warning=FALSE}
grid.arrange(fig_list$parameter_2100value$q10_Tgav, fig_list$parameter_2100value$q10_Ca)
```


\newpage

#### Fig 6.D: s
```{r, fig.width=6, fig.height=6, echo = FALSE, warning=FALSE}
grid.arrange(fig_list$parameter_2100value$s_Tgav, fig_list$parameter_2100value$s_Ca)
```

\newpage

### Fig 7: Parameter Space Bar Plots 
```{r, fig.width=8, fig.height=7.5, echo = FALSE, warning=FALSE}
fig_list$param_space$diff
```

```{r, fig.width=8, fig.height=7.5, echo = FALSE, warning=FALSE}
fig_list$param_space$s
```

```{r, fig.width=8, fig.height=7.5, echo = FALSE, warning=FALSE}
fig_list$param_space$q10
```

```{r, fig.width=8, fig.height=7.5, echo = FALSE, warning=FALSE}
fig_list$param_space$beta
```

