---
title: "Moving Average Years"
author: "K. Dorheim"
date: "March 1, 2018"
output: pdf_document
---

```{r, echo = F}

BASE <- "C:/Users/dorh012/Documents/GitHub/CMS/hector-SA-npar/methods_check"

co2_figs  <- get(load(file.path(BASE, "moving_average_years", "co2_figures.rda")))
flux_figs <- get(load(file.path(BASE, "moving_average_years", "landFlux_figures.rda")))

```



## CO2 

Comparison of the CO2 observation windows created by a 7 yr vs 15 year moving average. The observation windows really only differ at the beginning and end of the time series.

```{r, echo = F}
co2_figs$obs_ma
```


Table of counts for the number of runs that pass the observation windows created from the 7 and 15 year moving average. The windowing time does not really matter for the CO2 filtering method.

```{r, echo = F}
co2_figs$passing_count
```

\newpage

## Land Flux

Comparison of the land flux observation windows created by a 7 yr vs 15 year moving average. The 15 year moving average is not as extreme as the 7 year windowing method.

```{r, echo = F}
flux_figs$obs_ma
```

Table of counts for the number of runs that pass the observation windows created from the 7 and 15 year moving average. An extra 654 hector runs pass the the Land flux filtering method with the 15 year window (this is almost a 45% increase).

```{r, echo = F}
flux_figs$passing_count
```